### 2025-08-04 mixed-models.qmd
20250804 - Bayesian mixed model plots

# From GAM Model: Players at Top of End-of-Season Depth Chart
pointsPerSeason_positionAge_newData$fantasyPoints_depthGAM <- predict(
  object = pointsPerSeasonDepth_gam,
  newdata = pointsPerSeason_positionAge_newData,
  newdata.guaranteed = TRUE,
  exclude = "s(player_idFactor,ageCentered20)"
)
```

Plots of model-implied fantasy points by position and age are in Figures [-@fig-fantasyPointsByAgeQuadratic]–[-@fig-fantasyPointsByAgeGAMdepth].

@@

@@

>>>>

# From GAM Model: Players at Top of End-of-Season Depth Chart
pointsPerSeason_positionAge_newData$fantasyPoints_depthGAM <- predict(
  object = pointsPerSeasonDepth_gam,
  newdata = pointsPerSeason_positionAge_newData,
  newdata.guaranteed = TRUE,
  exclude = "s(player_idFactor,ageCentered20)"
)

# From Bayesian Model: All Players
model_levels <- levels(bayesianMixedModelFit$data$player_idFactor)

pointsPerSeason_positionAge_newData$player_idFactor <- factor(
  rep(levels(player_stats_seasonal_offense_subset$player_idFactor)[1], nrow(pointsPerSeason_positionAge_newData)), 
  levels = model_levels
)

modelPredictions_bayesian <- predict(
  object = bayesianMixedModelFit,
  newdata = pointsPerSeason_positionAge_newData,
  re_formula = NA,
  summary = TRUE
)

pointsPerSeason_positionAge_newData$fantasyPoints_bayesian <- modelPredictions_bayesian[,"Estimate"]
```

Plots of model-implied fantasy points by position and age are in Figures [-@fig-fantasyPointsByAgeQuadratic]–[-@fig-fantasyPointsByAgeBayesian].

@@

#### Bayesian Mixed Model {#sec-fantasyPointsByAgePlotsBayesian}

```{r}
#| label: fig-fantasyPointsByAgeBayesian
#| fig-cap: "Plot of Implied Trajectories of Fantasy Points by Age from a Bayesian Mixed Model."
#| fig-alt: "Plot of Implied Trajectories of Fantasy Points by Age from a Bayesian Mixed Model."
#| code-fold: true

ggplot2::ggplot(
  data = pointsPerSeason_positionAge_newData,
  mapping = aes(
    x = age,
    y = fantasyPoints_bayesian,
    color = positionFactor
  )
) + 
  geom_smooth() +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age and Position",
    subtitle = "Bayesian Mixed Model with All Players",
    color = "Position"
  ) +
  theme_classic() +
  guides(color = guide_legend(override.aes = list(fill = NA))) # transparent legend background
```

@@

We provide plots of individuals' model-implied fantasy points by age from the Bayesian mixed model in @sec-timeSeriesBayesianModelImpliedPredictions.
